---
title: "Distribution Transformation Explantion"
author: "Matthew Kay"
output: 
  html_document:
    self_contained: true
---

```{r knitr_setup, include=FALSE}
knitr::opts_chunk$set(
  fig.retina = 2,
  dev = "png",
  dev.args = list(png = list(type = "cairo"))
)
```

## Setup

The following libraries are needed to run this document:

```{r setup, message=FALSE}
library(ggplot2)
library(dplyr)
library(patchwork)
library(ggblend)   # remotes::install_github("mjskay/ggblend")
library(ggforce)
library(sn)
library(distributional)
library(ggdist)
library(purrr)
library(tidyr)

# theme_set(ggdist::theme_ggdist()) 
who <- toString(Sys.info())
```



```{r include=FALSE}

# Style setup

Settings <- list(
  Dot_Size_0 = 1.5,
  Dot_Size_1 = 2,
  Dot_Size_2 = 2.25,
  Dot_Size_3 = 3.5,

  Axis_Line_Size = 0.1,
  Font_Size_1 = 9,
  Stroke_Size = 0.5,
  
  Line_Size_0 = 0.25,
  Line_Size_1 = 0.5,
  Line_Size_2 = 0.65,
  Line_Size_3 = 1,
  Line_Size_4 = 1.25,
  
  Font_Size_0 = 6,
  Font_Size_1 = 8,
  Font_Size_2 = 10,
  
  Background_Color = "#FFFFFF",
  Text_Color = "#222222",
  Line_Color = "#222222",
  Line_Color_0 = "#888888",
  Line_Color_1 = "#999999",
  Line_Color_2 = "#cccccc",
  Axis_Color='#888888'
)


  Arrow_Setting = arrow (
      angle = 30,
      length = unit(0.05, "inches"),
      ends = "last",
      type = "open"
    )

HALF_EYE_SCALE = 1

Color_Dict <- list( graph = "#8F67A9",
                    text = "#68A175",
                    with_correction = '#77B0D1',
                    input = '#D5C168',
                    para = '#77B0D1', 
                    no_correction = "#aaaaaa",
                    graph_light = "#9f7daa",
                    text_light = "#65afa1",
                    dark_gradient = '#888888'
                   )

Blank_Color <- "#000000"

Color_Palette_Theory <- c(
  Color_Dict$no_correction,
  Color_Dict$with_correction
)


Color_Palette_Two <- c(
  Color_Dict$text,
  Color_Dict$graph
)

Color_Palette_Three<- c(
  Color_Dict$text,
  Color_Dict$graph,
  Color_Dict$no_correction
)

t <- theme(
                                      #line,
                                      #rect,
                                      #text,
  title = element_text(size = Settings$Font_Size_2, colour = Settings$Text_Color, face = "bold", vjust = 4.5), 
                                      #aspect.ratio,
  axis.title.x = element_text(size = Settings$Font_Size_1, colour = Settings$Text_Color, vjust = -2, hjust = 0.5), 
  axis.title.y = element_text(size = Settings$Font_Size_1,  colour = Settings$Text_Color, angle = 90, vjust = 2), 
                                      #?axis.title.x.top,
                                      #?axis.title.x.bottom,
  axis.text = element_text(size = Settings$Font_Size_1,  colour = Settings$Axis_Color, face = "plain"), 
                                      #?axis.title.y.left,
                                      #?axis.title.y.right,
  #axis.text.x,
                                      #?axis.text.x.top,
                                      #?axis.text.x.bottom,
  #axis.text.y,
                                      #?axis.text.y.left,
                                      #?axis.text.y.right,
  axis.ticks = element_line(colour = Settings$Axis_Color, size = Settings$Line_Size_1, linetype = "solid"),
  #axis.ticks.x,
                                      #?axis.ticks.x.top,
                                      #?axis.ticks.x.bottom,
  #axis.ticks.y,
                                      #?axis.ticks.y.left,
                                      #?axis.ticks.y.right,
  axis.ticks.length = unit(0, "lines"),
  #axis.ticks.length.x,
                                      #?axis.ticks.length.x.top,
                                      #?axis.ticks.length.x.bottom,
  axis.ticks.length.y = unit(0, "lines"),
                                      #?axis.ticks.length.y.left,
                                      #?axis.ticks.length.y.right,
  axis.line.x = element_line(colour = Settings$Axis_Color, size = Settings$Line_Size_1, linetype = "solid"),
  #axis.line.x,
                                      #?axis.line.x.top,
                                      #?axis.line.x.bottom,
  axis.line.y = element_line(colour = Settings$Axis_Color, size = Settings$Line_Size_1, linetype = "solid"),
                                      #?axis.line.y.left,
                                      #?axis.line.y.right,
  legend.background = element_blank(),
  legend.margin = margin(1, 1, 1, 1),
  legend.spacing = unit(0.5, "lines"),
                                      #?legend.spacing.x,
                                      #?legend.spacing.y,
  legend.key = element_blank(),
  legend.key.size = unit(1, "lines"),
  #legend.key.height,
  #legend.key.width,
  legend.text = element_text(size = Settings$Font_Size_0,  colour = Settings$Text_Color),
  legend.text.align = 0,#left
  legend.title = element_text(size = Settings$Font_Size_0, colour = Settings$Text_Color, vjust = 0.5),
  legend.title.align = 0,#left
  legend.position = "none",
  #legend.direction, #auto
  legend.justification = "center",
  #legend.box = element_rect(fill = NA, colour = Line_Color_0, size = 0.1), #auto
  legend.box.just = "left", 
  legend.box.margin = margin(1, 1, 1, 1),
  legend.box.background = element_blank(),
  legend.box.spacing = unit(1, "lines"),
  panel.background = element_blank(),
  panel.border = element_blank(),
  panel.spacing = unit(2, "lines"),
  #panel.spacing.x,
  #panel.spacing.y,
  #panel.grid,
  panel.grid.major.y = element_blank(),#element_line(colour = Settings$Line_Color_2, size = Settings$Line_Size_0, linetype = "solid"),
  panel.grid.major.x = element_blank(),
  panel.grid.minor = element_blank(),
  #panel.grid.major.y = element_line(colour = Line_Color_2, size = Line_Size_0, linetype = "solid"),
  #panel.grid.major.y,
  #panel.grid.minor.x,
  #panel.grid.minor.y, 
  panel.ontop = FALSE,
  plot.background = element_blank(),
  plot.title  = element_text(size = Settings$Font_Size_2,  colour = Settings$Text_Color, hjust = 0.5, vjust = 3),
  #?plot.title.position,
  plot.subtitle  = element_text(size = Settings$Font_Size_1,  colour = Settings$Text_Color, hjust = 0.5, vjust = .5),
  plot.caption = element_text(size = Settings$Font_Size_2,  colour = Settings$Text_Color, hjust = 0.5, vjust = .5),
  #?plot.caption.position,
  #?plot.tag,
  plot.tag.position = "top",
  plot.margin = margin(10, 10, 10, 10),
  strip.background = element_blank(),
  #strip.background.x,
  #strip.background.y,
  strip.placement = "outside",
  strip.text = element_text(size = Settings$Font_Size_2,  face = "bold", colour = Settings$Text_Color, hjust = 0.5, vjust = .5),
  #strip.text.x,
  #strip.text.y,
  #strip.switch.pad.grid,
  #strip.switch.pad.wrap,
  complete = FALSE,
  validate = TRUE
) 


theme_set(t)
```

```{r}
# quartzFonts()
ggplot <-
  function(...) ggplot2::ggplot(...) +
  scale_color_manual(values = Color_Palette_Theory) + 
  scale_fill_manual(values = Color_Palette_Theory) + 
  scale_x_continuous(breaks = seq(0, 1, by = .2), limits = c(0, 1), expand = c(0.01, 0.01)) + 
  scale_y_continuous(breaks = seq(0, 1, by = .2), limits = c(0, 1), expand = c(0.01, 0.01))


set.seed(1001)

prob_seq = seq(.01, 1.00, length.out = 25)
```


## The linear-in-probit model of perceived probabilities

We want to set up a linear-in-probit function that maps from
true probabilities to perceived probabilities in probit space. It has two parameters:

- `alpha` ($\alpha$): the intercept: the perceived probability at 50%, in probit space
- `beta` ($\beta$): the slope of the linear relationship between true and perceived
probabilities, in probit space.

The linear-in-probit function for a perceived probability, $p_\mathrm{perc}$, 
given a displayed probability, $p$, is:

$$
p_\mathrm{perc} 
= \mathrm{lpr}(p)
= \mathrm{probit}^{-1}\left( \alpha + \beta \cdot \mathrm{probit}(p) \right)
$$

To write this in R, we need the `probit` and `inv_probit` functions, which are
the inverse CDF and CDF of the normal distribution, respectively:

```{r}
probit = qnorm
inv_probit = pnorm
```

Then we can write the `lpr` function and its inverse as:

```{r}
lpr = function(p, alpha, beta) inv_probit(alpha + beta * probit(p))
inv_lpr = function(p, alpha, beta) inv_probit((probit(p) - alpha) / beta)
```

This function looks like this:

```{r fig.width = 5, fig.height = 7}
lpr_grid = expand.grid(
  p = seq(0, 1, length.out = 200),
  # F: I pull put the curves so that people can see them
  alpha = c(-.8, 0, .8),
  beta = exp(seq(-.8, .8, length.out = 6))
) |>
  mutate(p_perc = lpr(p, alpha, beta))


plot_lpr_grid_probit = lpr_grid |>
  filter(0 < p & p < 1) |>
  ggplot(aes(x = probit(p), y = probit(p_perc), group = interaction(alpha, beta), color = beta)) +
  #geom_vline(xintercept = 0, color = Settings$Line_Color) +
  #geom_hline(yintercept = 0, color = Settings$Line_Color) +
  #geom_abline(intercept = 0, slope = 1, color = Settings$Line_Color) +
  # F: I use segment for convenience in post hoc edit
  geom_segment(x = -2.59, xend = 2.59, y = -2.59, yend = 2.59, alpha = .25, linetype = "22", color = 'black') +
  scale_x_continuous() + 
  scale_y_continuous() + 
  scale_color_gradient(low = Color_Dict$dark_gradient, high = Color_Dict$graph) + 
  #scale_fill_gradient(low = Color_Dict$dark_gradient, high = Color_Dict$para) + 
  labs(subtitle = "lpr in probit space", y = "probit(subjective probability)", x = "probit(true probability)")


plot_lpr_grid = lpr_grid |>
  ggplot(aes(x = p, y = p_perc, group = interaction(alpha, beta), color = beta)) +
  #geom_vline(xintercept = 0.5, color = Settings$Line_Color) +
  #geom_hline(yintercept = 0.5, color = Settings$Line_Color) +
  geom_segment(x = 0, xend = 1, y = 0, yend = 1, alpha = 0.25, size = 0.75, linetype = "22", color = 'black') +
  scale_color_gradient(low = Color_Dict$dark_gradient, high = Color_Dict$graph) + 
  #scale_fill_gradient(low = Color_Dict$dark_gradient, high = Color_Dict$input) + 
  labs(subtitle = "lpr in probability space", y = "subjective probability", x = "true probability")


(plot_lpr_grid_probit + plot_lpr_grid) &
  geom_line(show.legend = FALSE, size = 1) &
  facet_wrap(vars(alpha), ncol = 1, labeller = "label_both", scales = 'free') #&
  #cowplot::panel_border()
```

```{r include=FALSE}


if(str_detect(who, 'fm')){ # if this is fm
# sorry I don't know how to adjust widths in ggblend....
combined <- ggarrange(plot_lpr_grid_probit +
          geom_line(show.legend = FALSE, size = 1)  +
          facet_wrap(vars(alpha), ncol = 1, labeller = "label_both", scales = 'free') , 
          plot_lpr_grid + 
          geom_line(show.legend = FALSE, size = 1)  +
          facet_wrap(vars(alpha), ncol = 1, labeller = "label_both", scales = 'free'),
          ncol = 2, widths = c(2, 3))

   ggsave(combined, file = "~/Documents/Github/Bias-correcion-paper/fig/linear-in-probit-f.pdf", width = 350, height = 650, scale = 4, units = 'px')
} else { 
   ggsave("figures/linear-in-probit.pdf", width = 5, height = 5)
}
```


## Transforming distributions through the linear-in-probit model

Given a distribution we want to display:

$$
X \sim \mathrm{Normal}(\mu, \sigma)
$$

We can use the perceptual transformation using the linear-in-probit function
from [Kay 2019](https://osf.io/6xcnw/) to determine a distribution we should
display that will be perceived as if it were $X$, from the perspective of
judging the probability that distribution exceeds a given threshold; call this $X^>$:

$$
\begin{align*}
\mathbb{P}(X^> > x) &= \mathrm{lpr}^{-1}\left( \mathbb{P}(X > x) \right)\\
\implies X^> &\sim \mathrm{Normal}(\mu^>, \sigma^>)\\
\textrm{where } \mu^> &= \mu - \alpha\sigma\\
\textrm{and } \sigma^> &= \beta\sigma
\end{align*}
$$

First, let's define the `alpha` and `beta` parameters of the linear-in-probit
probability perception function we'll use for the rest of this document:

```{r}
alpha = -.34
beta = 2.44
```

If we have a distribution we want to display, say $X \sim Normal(\mu, \sigma)$:

```{r}
mu = .468
sigma = .025

X_perceived = dist_normal(mu, sigma)
X_perceived
```

We can define a function for $X^>_{\alpha,\beta}$, also parameterized by $\mu$ and $\sigma$:

```{r}
dist_gt_normal = function(mu, sigma) {
  mu_gt = mu - alpha * sigma
  sigma_gt = beta * sigma
  dist_normal(mu_gt, sigma_gt)  
}

X_gt = dist_gt_normal(mu, sigma)
X_gt
```

### Visualizing the transformation

To visualize the transformation, we will build up a set of subplots that
we can arrange into a larger plot show the relationship between the percevied
distribution, the linear-in-probit function, and the displayed distribution.

**NOTE:** these pieces may not make sense in isolation; you may want to skip
to **Putting it all together**, below.

#### Densities

We'll start by visualizing both distributions together as densities:

```{r}
x_min = 0.3
x_max = 0.7

arrow_ = \(...) arrow(..., angle = 45, type = "closed", length = unit(5, "pt"))

geom_segment_ = \(..., alpha = 0.5, linetype = "11") geom_segment(..., alpha = alpha, linetype = linetype)

# scale_color_dist = \(...) scale_color_brewer(..., 
#   limits = c("perceived", "displayed"), palette = "Set2",
#   guide = "none",
#   aesthetics = c("color", "fill")
# )

scale_color_dist = \(...) scale_color_manual(..., 
  values = c("perceived" = Color_Dict$no_correction, "displayed" = Color_Dict$para),
  guide = "none",
  aesthetics = c("color", "fill")
)

quantile_lines_df = function(X_perceived, X_displayed) {
  tibble(
    x_perc = seq(
      0.45,
      0.6,
      by = 0.025
    ),
    p_perc = 1 - cdf(X_perceived, x_perc)[[1]],
    d_perc = density(X_perceived, x_perc)[[1]],
    p_disp = inv_lpr(p_perc, alpha, beta),
    x_disp = quantile(X_displayed, 1 - p_disp)[[1]],
    d_disp = density(X_displayed, x_disp)[[1]]
  )
}

density_plot = function(X_perceived, X_displayed) {
  quantile_lines_df = quantile_lines_df(X_perceived, X_displayed)
  
  data.frame(
    X = c(X_perceived, X_displayed),
    dist = c("perceived", "displayed")
  ) |>
    ggplot() +
    stat_slab(
      aes(xdist = X, color = dist, fill = dist, side = ifelse(dist == "perceived", "top", "bottom")), 
      alpha = 0.5, size = 0.75
    ) +
    geom_segment_(
      aes(x = x_perc, xend = x_perc, y = 0, yend = 1, color = "perceived"),
      data = quantile_lines_df
    ) +
    geom_segment_(
      aes(x = x_disp, xend = x_disp, y = 0, yend = -1, color = "displayed"),
      data = quantile_lines_df
    ) +
    geom_segment_(
      aes(x = x_perc, xend = x_perc, y = 0, yend = 1, color = "perceived"),
      data = filter(quantile_lines_df, x_perc == 0.5), linetype = "solid", alpha = 1, size = 1.25
    ) +
    geom_segment_(
      aes(x = x_disp, xend = x_disp, y = 0, yend = -1, color = "displayed"),
      data = filter(quantile_lines_df, x_perc == 0.5), linetype = "solid", alpha = 1, size = 1.25
    ) +
    coord_cartesian(expand = FALSE, clip = "off", xlim = c(x_min, x_max)) +
    scale_color_dist() + 
    scale_x_continuous(limits = c(x_min, x_max), breaks = NULL) +
    scale_y_continuous(breaks = NULL) +
    labs(x = NULL, y = NULL)  +
     theme(
       axis.line.x = element_blank(),
       axis.line.y = element_blank()#,
    #   plot.background = element_blank(),
    #   panel.background = element_blank()
     )    
}

density_plot(X_perceived, X_gt)
```

#### CCDF of perceived distribution

If we look at the CCDF of the desired-to-be-perceived distribution:

```{r}
perceived_ccdf_plot = function(X_perceived, X_displayed) {
  quantile_lines_df = quantile_lines_df(X_perceived, X_displayed)
  
  data.frame(
    X = X_perceived
  ) |>
    ggplot(aes(color = "perceived")) +
    stat_slab(aes(xdist = X), slab_type = "ccdf", normalize = "none", fill = NA, size = 1.25, scale = 1) +
    geom_segment_(
      aes(x = x_perc, xend = x_perc, y = 0, yend = p_perc, color = "perceived"),
      data = quantile_lines_df
    ) * copy_over(
      aes(xend = x_max, y = p_perc)
    ) +
    geom_segment_(
      aes(x = x_perc, xend = x_perc, y = 0, yend = p_perc, color = "perceived"), size = 1.25,
      data = filter(quantile_lines_df, x_perc == 0.5), alpha = 1, linetype = "solid"
    ) * copy_over(
      aes(xend = x_max, y = p_perc),
      arrow = arrow_()
    ) +
    scale_y_continuous(limits = c(0,1)) +
    scale_x_continuous(limits = c(x_min, x_max), breaks = NULL) +
    scale_color_dist() +
    coord_cartesian(expand = FALSE, clip = "off", xlim = c(x_min, x_max)) +
    labs(
      x = NULL,
      y = "Perceived P(X > x)"
    )  +
     theme(
       #plot.background = element_blank(),
       #panel.background = element_blank(),
       axis.line.x = element_blank()
     )
}

perceived_ccdf_plot(X_perceived, X_gt)
```

#### Linear-in-probit function

We can translate perceived probabilities into corresponding probabilities we should display:

```{r}
lpr_plot = function(X_perceived, X_displayed) {
  quantile_lines_df = quantile_lines_df(X_perceived, X_displayed)
  
  lpr_df = tibble(
    p_disp = seq(0, 1, length.out = 500),
    p_perc = lpr(p_disp, alpha, beta)
  ) 
  
  lpr_df |>
    ggplot(aes(p_disp, p_perc)) +
    geom_line(size = 1.25, color = Color_Dict$graph) +
    geom_segment_(
      aes(x = 0, xend = p_disp, y = p_perc, yend = p_perc, color = "perceived"), 
      data = quantile_lines_df
    ) * copy_over(
      aes(x = p_disp, y = 0, color = "displayed")
    ) +
    geom_segment_(
      aes(x = 0, xend = p_disp, y = p_perc, yend = p_perc, color = "perceived"), size = 1.25,
      data = filter(quantile_lines_df, x_perc == 0.5), alpha = 1, linetype = "solid"
    ) * copy_over(
      aes(x = p_disp, y = 0, color = "displayed")
    ) +
    geom_abline(intercept = 0, slope = 1, alpha = 0.25, size = 0.75, linetype = "22") +
    scale_x_continuous(limits = c(0, 1), position = "top") +
    scale_y_continuous(limits = c(0, 1), position = "right") +
    scale_color_dist() +
    coord_cartesian(expand = FALSE, clip = "off") +
    labs(
      x = "Displayed P(X > x)",
      y = "Perceived P(X > x)"
    ) +
    theme(
      #plot.background = element_blank(),
      panel.background = element_rect(fill = "gray95", color = alpha("black", 0.25)),
    )
}

lpr_plot(X_perceived, X_gt)
```

#### Connecting plots

We will also need two plots that are just connectors to draw lines and arcs from the
linear-in-probit plot to the CCDF of the displayed distribution.

First, a plot of lines:

```{r}
lines_plot = function(X_perceived, X_displayed) {
  quantile_lines_df = quantile_lines_df(X_perceived, X_displayed)

  quantile_lines_df |>
    ggplot(aes(x = p_disp, xend = p_disp, y = 0, yend = 1, color = "displayed")) +
    geom_segment_() +
    geom_segment_(data = \(d) d |> filter(x_perc == 0.5), linetype = "solid", alpha = 1,  size = 1.25) +
    scale_color_dist() +
    coord_cartesian(expand = FALSE, clip = "off") +
    labs(x = NULL, y = NULL) +
    scale_x_continuous(limits = c(0, 1), breaks = NULL) +
    scale_y_continuous(limits = c(0, 1), breaks = NULL) +
    theme(
       axis.line.x = element_blank(),
       axis.line.y = element_blank()#,
    #   plot.background = element_blank(),
    #   panel.background = element_blank()
     )
}

lines_plot(X_perceived, X_gt)
```

And a plot of curves:

```{r}
curves_plot = function(X_perceived, X_displayed) {
  quantile_lines_df = quantile_lines_df(X_perceived, X_displayed)

  quantile_lines_df |>
    ggplot(aes(x0 = 0, y0 = 0, r = p_disp, start = pi/2, end = pi, color = "displayed")) +
    geom_arc(alpha = 0.5, linetype = "11") +
    geom_arc(data = \(d) d |> filter(x_perc == 0.5), arrow = arrow_(),  size = 1.25) +
    scale_x_continuous(limits = c(0,1), breaks = NULL) +
    scale_y_reverse(limits = c(1,0), breaks = NULL) +
    coord_cartesian(expand = FALSE, clip = "off") +
    scale_color_dist() +
    labs(x = NULL, y = NULL) +
    theme(
       axis.line.x = element_blank(),
       axis.line.y = element_blank() #,
    #   plot.background = element_blank(),
    #   panel.background = element_blank()
    )
}

curves_plot(X_perceived, X_gt)
```

#### CCDF of displayed distribution

Finally, we need the CCDF of the displayed distribution, flipped to line up
with the flipped density in the density plot:

```{r}
displayed_ccdf_plot = function(X_perceived, X_displayed) {
  quantile_lines_df = quantile_lines_df(X_perceived, X_displayed)

  data.frame(
    X = X_displayed
  ) |>
    ggplot(aes(color = "displayed")) +
    stat_slab(aes(xdist = X), slab_type = "ccdf", normalize = "none", fill = NA, size = 1.25, scale = 1, side = "bottom") +
    geom_segment_(
      aes(x = x_disp, xend = x_disp, y = 0, yend = p_disp, color = "displayed"),
      data = quantile_lines_df
    ) * copy_over(
      aes(xend = x_max, y = p_disp)
    ) +
    geom_segment_(
      aes(x = x_disp, xend = x_disp, y = 0, yend = p_disp, color = "displayed"),  size = 1.25,
      data = filter(quantile_lines_df, x_perc == 0.5), linetype = "solid", alpha = 1
    ) * copy_over(
      aes(xend = x_max, y = p_disp),  size = 1.25
    ) +
    scale_color_dist() +
    coord_cartesian(expand = FALSE, clip = "off", xlim = c(x_min, x_max)) +
    labs(
      x = "x",
      y = "Displayed P(X > x)"
    ) +
    scale_y_reverse(limits = c(1,0)) +
    scale_x_continuous(limits = c(x_min, x_max)) #+
    # theme(
    #   plot.background = element_blank(),
    #   panel.background = element_blank()
    # )
}

displayed_ccdf_plot(X_perceived, X_gt)
```

#### Putting it all together

Putting the pieces together above we have:

```{r fig.width = 6, fig.height = 6}
transformation_plot = function(X_perceived, X_displayed) {
  (
    (perceived_ccdf_plot(X_perceived, X_displayed) + lpr_plot(X_perceived, X_displayed)) /
    (density_plot(X_perceived, X_displayed) + lines_plot(X_perceived, X_displayed)) /
    (displayed_ccdf_plot(X_perceived, X_displayed) + curves_plot(X_perceived, X_displayed))
  ) & 
    theme(plot.margin = margin())
}

transformation_plot(X_perceived, X_gt)
```

```{r}
if(str_detect(who, 'fm')){ # if this is fm
  ggsave("~/Documents/Github/Bias-correcion-paper/fig/correction-loop-normal-f.pdf", width = 400, height = 550, scale = 5, units = 'px')
} else { 
  ggsave("figures/correction-loop-normal.pdf", width = 6, height = 6)
}

```


If you trace a single line clockwise around this chart, it shows:
1. We can take an x position in the green distribution (the probabilities we want to be perceived,
2. Take the probability that X > x,
3. Transform it according to the inverse linear-in-probit to obtain the probability we should show,
4. and then look up the x value corresponding to that probability in $X^>$ (the orange distribution),
and find that it is the same x value we started with. This holds for all $x$ values. Thus, $X^>$ 
is a distribution that satisfies the property that *all* of the right-tailed probabilities it shows should
be perceived as the right-tailed probabilities from $X$ (the green distribution).

## Skew-normal

As shown above, $X^>$ ensures that *every* right-tailed probability that is 
displayed is perceived as the corresponding right-tailed probability on the underlying
distribution. However, in order to do so, it must also move the mode of the distribution,
which may be a bit weird --- especially if the mode is intended to match the most likely
vote count predicted by the model.

We could try to find a different distribution, say $X^*$, where:

1. The mode of the displayed distribution matches the mode of the
underlying distribution

2. $P(X^* > 0.5)$ equals $\mathrm{lpr}^{-1}(P(X > 0.5))$, since after all this is the win probability --- 
the probability we care most about.

3. For the remaining $x$ values, $P(X^* > x)$ *roughly* equals $P(X > x)$.

We'll do this by finding a skew-normal with mode equal to $mu$, standard deviation 
equal to $sigma^>$, and $P(X^* > 0.5)$ = $P(X > 0.5)$.

To do this, we first define a skew-normal in terms of its sd (`sigma`) instead of
the usual `omega` scale parameter (the usual parameterization is in terms of location, `xi`, 
scale, `omega`, and skew, `alpha`). This alternative parameterization
will allow us to generate a distribution with a specific standard deviation:

```{r}
omega_ = function(sigma, alpha) {
  delta = alpha / sqrt(1 + alpha^2)
  sigma / sqrt(1 - 2 * delta^2 / pi)
}

dsn_sigma = function(x, xi, sigma, alpha) {
  dsn(x, xi, omega_(sigma, alpha), alpha)  
}

qsn_sigma = function(p, xi, sigma, alpha) {
  qsn(p, xi, omega_(sigma, alpha), alpha)  
}

psn_sigma = function(x, xi, sigma, alpha) {
  psn(x, xi, omega_(sigma, alpha), alpha)  
}

rsn_sigma = function(n, xi, sigma, alpha) {
  rsn(n, xi, omega_(sigma, alpha), alpha)  
}

```

Next, we need a function to get the mode of a skew Normal distribution. The mode
of a skew-Normal does not have a closed form, so we
will use numerical optimization to find the maximum of the density function:

```{r}
sn_mode = function(xi = 0, sigma = 1, alpha = 0) {
  optim(
    qsn_sigma(0.5, xi, sigma, alpha), # start at median
    dsn_sigma, 
    xi = xi, sigma = sigma, alpha = alpha, 
    control = list(fnscale = -1),     # maximize
    method = "BFGS"
  )$par
}
```

Now we can find a skew-normal distribution with approximately a mode equal to
the original mean (`mu`), standard deviation equal to the lpr-transformed distribution
(`sigma_gt`), and a `P(X > 0.5)` equal to the probability
for X being greater than 0.5 that we want to display (which will be perceived
as the value we want):

```{r}
# vectorize this so we can use it later to construct vectors of distributions easily
find_skew_normal_params = Vectorize(function(mu, sigma) {
  mu_gt = mu - alpha * sigma
  sigma_gt = beta * sigma
  
  desired_mode = mu
  desired_sigma = sigma_gt
  desired_x = 0.5
  desired_p_gt_x = inv_lpr(pnorm(desired_x, mu, sigma, lower.tail = FALSE), alpha, beta)
  
  starting_xi = mu_gt
  # start alpha not at 0 to avoid a local minimum
  starting_alpha = if (desired_mode < mu_gt) 1 else -1
  
  cost = function(x) {
    xi = x[[1]]
    alpha = x[[2]]
    # since both costs here are squares of differences of percentages, multiply 
    # by 100^2 to get cost function on a reasonable scale
    10000 * (
      (sn_mode(xi, desired_sigma, alpha) - desired_mode)^2 +
      (1 - psn_sigma(desired_x, xi, desired_sigma, alpha) - desired_p_gt_x)^2
    )
  }
  
  params = optim(c(xi = starting_xi, alpha = starting_alpha), cost)$par
  c(params["xi"], sigma = desired_sigma, params["alpha"])
})

params = find_skew_normal_params(mu, sigma)
params
```

Let's check:

```{r}
desired_x = 0.5
desired_p_gt_x = inv_lpr(pnorm(.5, mu, sigma, lower.tail = FALSE), alpha, beta)

cat(sep = "",
  "Desired mode: ", mu, "\n",
  "Actual mode:  ", sn_mode(params["xi",], params["sigma",], params["alpha",]), "\n",
  "Desired P(X > ", desired_x, "): ", desired_p_gt_x, "\n",
  "Actual P(X > ", desired_x, "):  ", 1 - psn_sigma(desired_x, params["xi",], params["sigma",], params["alpha",]), "\n"
)
```

We can put this together to create a function that generates vectors of
skew-normal distributions matching the above constraints:

```{r}
dist_gt_skew_normal = function(mu, sigma) {
  params = find_skew_normal_params(mu, sigma)
  dist_wrap("sn_sigma", params["xi",], params["sigma",], params["alpha",])
}

X_gt_skew = dist_gt_skew_normal(mu, sigma)
X_gt_skew
```

The transformation now looks like this:

```{r fig.width = 6, fig.height = 6}
transformation_plot(X_perceived, X_gt_skew)
```



We can see that at x = 0.5, the right-tailed probability matches. The other quantiles
roughly match, though not as well as with $X^>$. However, the *modes* of the two
distributions are equal here, unlike with $X^>$.


```{r}
if(str_detect(who, 'fm')){ # if this is fm
   ggsave("~/Documents/Github/Bias-correcion-paper/fig/correction-loop-skew-f.pdf", width = 400, height = 550, scale = 5, units = 'px')
} else { 
  ggsave("figures/correction-loop-skew.pdf", width = 6, height = 6)
}
```

### Comparing the three approaches

Comparing the three approaches on one value of mu:

```{r}
data.frame(
  dist = c(
    "uncorrected",
    "corrected normal",
    "corrected skew-normal"
  ),
  X = c(
    dist_normal(mu, sigma),
    dist_gt_normal(mu, sigma),
    dist_gt_skew_normal(mu, sigma)
  )
) |>
  ggplot(aes(xdist = X, color = dist)) +
  stat_slab(fill = NA) |> blend("darken") +
  geom_vline(xintercept = mu, alpha = 0.5) +
  scale_color_brewer(palette = "Set2")
```

We can see that the modes match for the skewed distribution, and it is not too far
off the shape of the fully-corrected normal distributoin.

Let's check over a set of distributions with different `mu`:

```{r}
dists = c(
  "uncorrected" = dist_normal,
  "corrected normal" = dist_gt_normal,
  "corrected skew-normal" = dist_gt_skew_normal
)

dist_df = map_dfr(dists, .id = "correction", \(d) {
  tibble(
    mu = seq(0.47, 0.53, length.out = 7),
    dist = d(mu, sigma)
  )
})

dist_df
```

Let's check that the modes are indeed correct by looking at the difference
between the modes of the distributions and the desired `mu`:

```{r}
dist_df |>
  mutate(mode = round(ggdist::Mode(dist) - mu, 3)) |>
  select(correction, mode, mu) |>
  pivot_wider(names_from = correction, values_from = mode)
```

We can see the modes for the corrected normal distribution are all above the desired
mode (as expected, due to a negative `alpha` parameter in the linear-in-probit function),
while the corrected skew-normal distribution has a mode roughly equal to `mu`.


```{r fig.width = 11, fig.height = 8}
dist_df |>
  ggplot(aes(xdist = dist, color = correction)) +
  stat_slab(fill = NA) |> blend("darken") +
  scale_color_brewer(palette = "Set2") +
  facet_wrap(vars(mu)) +
  geom_vline(aes(xintercept = mu), alpha = 0.5) 
```

The skew-normal corrected version is indeed very similar to the normal corrected version, but
has its mode at the input value of `mu`.

